---
title: "From PROC COMPARE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{From PROC COMPARE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ksCompare)
```

This article maps SAS `PROC COMPARE` syntax onto `ksCompare`. The goal is
spirit-equivalence, not byte-identical output: `ksCompare` returns
tibbles you can render however you like.

## Cheat sheet

| SAS                                 | ksCompare                                                 |
|-------------------------------------|-----------------------------------------------------------|
| `BASE = ds1`                        | `base = ds1`                                              |
| `COMPARE = ds2`                     | `comp = ds2`                                              |
| `ID v1 v2`                          | `by = c("v1", "v2")`                                      |
| `BY v1 v2`                          | `by = c("v1", "v2")` (also covers BY-group reporting)     |
| `VAR x y`                           | `mapping = c(x = "x", y = "y")`                           |
| `WITH x2`                           | `mapping = c(x = "x2")`                                   |
| `CRITERION = 1e-5`                  | `tolerance = ks_tol(abs = 1e-5)`                          |
| `METHOD = RELATIVE` / `PERCENT`     | `tolerance = ks_tol(rel = 1e-5)`                          |
| `METHOD = EXACT`                    | `tolerance = ks_tol()` (default, strict)                  |
| (floating-point round-trip noise)   | `tolerance = ks_tol(ulp = 4)`                             |
| `NOMISS`                            | `options = ks_options(na_equal = FALSE)`                  |
| `OUT=`                              | `as_outbase()` / `as_outcomp()`                           |
| `OUTDIF`                            | `as_outdif()`                                             |
| `OUTNOEQUAL`                        | `as_outnoequal()`                                         |
| `&SYSINFO` bitmask                  | `ks_sysinfo()`                                            |
| `NOPRINT`                           | use the returned object directly; no auto-print           |
| `LISTALL` / `LISTBASE` / `LISTCOMP` | rows are tagged in `cmp$row_diff$status`                  |

## Row matching strategies

| Situation                                  | ksCompare argument                  |
|--------------------------------------------|-------------------------------------|
| No keys, match by row position             | `by = NULL` (default)               |
| Same key column on both sides              | `by = "USUBJID"`                    |
| Renamed key column                         | `by = c(USUBJID = "SUBJID")`        |
| Discover the key automatically             | `by = "auto"`                       |
| Duplicate keys: keep first occurrence      | `dup_keys = "first"` (default)      |
| Duplicate keys: keep last occurrence       | `dup_keys = "last"`                 |
| Duplicate keys: pair within group (1<->1)  | `dup_keys = "keep_all"`             |
| Duplicate keys: cartesian product          | `dup_keys = "all_pairs"`            |
| Duplicate keys: hard fail                  | `dup_keys = "error"`                |

## OUT* outputs

```{r}
a <- data.frame(id = 1:3, x = c(1, 2, 3))
b <- data.frame(id = 1:3, x = c(1, 2, 4))
cmp <- ks_compare(a, b, by = "id")

as_outbase(cmp)
as_outcomp(cmp)
as_outdif(cmp)
as_outnoequal(cmp)
```

## SYSINFO bitmask

```{r}
si <- ks_sysinfo(cmp)
si
attr(si, "bits")
```

The bit names follow the SAS documentation; only flags with a clear
correspondence in tibble-land are populated. Extra ksCompare findings
(patterns, suggestions) live on the `ks_comparison` object directly.

## Things ksCompare does not try to replicate

- The fixed-width LISTING-style `PROC COMPARE` printout. Use
  `print(cmp)`, `summary(cmp)`, or `ks_report_html(cmp)`.
- `OUT=` data sets in PDV order. Output tibbles use a clean
  long/wide layout that is easier to filter and join.
