---
title: "Getting started with ksCompare"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ksCompare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ksCompare)
```

`ksCompare` compares two data frames in the spirit of SAS `PROC COMPARE` —
column-by-column, row-by-row, with tolerances and labelled-metadata
awareness — and returns a structured `ks_comparison` object you can
print, query, export, or render to HTML / Excel.

## A first comparison

```{r}
base <- data.frame(
  id   = 1:5,
  age  = c(34, 41, 28, 55, 19),
  arm  = c("A", "A", "B", "B", "A"),
  stringsAsFactors = FALSE
)
comp <- data.frame(
  id   = 1:5,
  age  = c(34, 41, 27, 55, 19),
  arm  = c("A", "A", "b", "B", "A"),
  stringsAsFactors = FALSE
)

cmp <- ks_compare(base, comp, by = "id")
cmp
```

`summary(cmp)` returns a structured object with per-section counts;
`as_tibble(cmp)` returns the long-format value-diff table.

```{r}
summary(cmp)
tibble::as_tibble(cmp) |> head()
```

## What's inside `cmp`?

A `ks_comparison` is a list with named slots; you can read them
directly when scripting QC dashboards:

```{r}
names(cmp)
```

- `cmp$schema_diff` -- one row per column pair (matched, base-only,
  or comp-only) with `kind_match`, `label_match`, `format_match`.
- `cmp$row_diff` -- one row per matched / unmatched row with
  `key_id`, `base_row`, `comp_row`, `status`.
- `cmp$value_diff` -- one row per differing cell with `kind`, `base`,
  `comp`, `diff`, `note`. Same as `as_tibble(cmp)` and `ks_tidy(cmp)`.
- `cmp$pattern_summary` -- recurring shapes per column (e.g.
  `case_only`, `constant_offset`).
- `cmp$column_suggestions` -- fuzzy rename candidates (when
  `stringdist` is installed).
- `cmp$meta` -- counts, key columns, and the row-matching strategy.
- `cmp$manifest` -- input hashes + run timestamp for QC traceability.

```{r}
cmp$schema_diff
cmp$value_diff
ks_glance(cmp)
```

## Tolerances

```{r}
ks_compare(
  data.frame(id = 1, x = 1.0),
  data.frame(id = 1, x = 1.0001),
  by = "id",
  tolerance = ks_tol(abs = 1e-3)
)
```

Per-column overrides:

```{r, eval = FALSE}
ks_tol(
  abs = 0,
  per_column = list(
    weight = ks_tol(abs = 0.01),
    height = ks_tol(rel = 0.001)
  )
)
```

ULP tolerance is useful when you only want to absorb floating-point
round-trip noise:

```{r, eval = FALSE}
ks_compare(base, comp, by = "id", tolerance = ks_tol(ulp = 4))
```

## Options

`ks_options()` controls semantics that apply to all columns:

- `na_equal`: treat `NA == NA` as equal (default `TRUE`).
- `sas_special_missing`: distinguish `.A`-`.Z` and `._` SAS missings
  (haven-imported numerics only).
- `compare_labels` / `compare_formats`: schema-level checks. Format
  comparison is trailing-dot- and case-tolerant
  (`DATE9.` == `DATE9` == `date9.`).
- `str_trim`, `str_case`, `str_norm`: string normalisation. Strings
  are always compared encoding-safely (Latin1 vs UTF-8 will not
  produce false diffs).
- `tz`: `"preserve"` (default), `"UTC"`, or `"strip"` for `POSIXct`.

```{r, eval = FALSE}
ks_compare(
  base, comp, by = "id",
  options = ks_options(str_trim = TRUE, str_case = "fold")
)
```

## Renamed columns

Use `mapping =` for renamed non-key columns and a named `by =` for a
renamed key:

```{r, eval = FALSE}
ks_compare(
  base, comp,
  by      = c(USUBJID = "SUBJID"),
  mapping = c(AGE = "AGE_YEARS")
)
```

## Reading files directly

`ks_compare()` accepts file paths in addition to in-memory data frames
and dispatches the right reader by extension:

```{r, eval = FALSE}
# SAS transport file
ks_compare("prod/adsl.xpt", "qc/adsl.xpt", by = "USUBJID")

# .sas7bdat
ks_compare("prod/adsl.sas7bdat", "qc/adsl.sas7bdat", by = "USUBJID")

# Parquet / Feather
ks_compare("a.parquet", "b.parquet", by = "id")

# RDS / RData / CSV / TSV are also recognised
ks_compare("a.rds", "b.rds", by = "id")
```

When both paths share the same basename (`prod/adsl.xpt` vs
`qc/adsl.xpt`), display names are automatically disambiguated using
the parent folder, so `print(cmp)` and the HTML report show
`prod/adsl` vs `qc/adsl` instead of two identical labels.

## Duplicate keys

By default `dup_keys = "first"` keeps the first occurrence per side
(emitting a `ksCompare_dup_keys_resolved` info message). Other
strategies:

- `"last"` -- keep the last per side.
- `"keep_all"` -- pair duplicates positionally within each key.
- `"all_pairs"` -- cartesian-pair duplicates (warns when group sizes
  differ).
- `"error"` -- raise on any duplicate key.

```{r, eval = FALSE}
ks_compare(base, comp, by = "id", dup_keys = "keep_all")
```

## Next steps

- **From SAS:** see `vignette("from-proc-compare")` for a side-by-side
  mapping of `PROC COMPARE` options.
- **Smart features:** see `vignette("smart-features")` for auto-key
  inference, fuzzy column suggestions, and pattern detection.
- **Reports:** see `vignette("reports")` for HTML and Excel output.
- **CI / pipelines:** see `vignette("pipeline-gates")` for using
  comparisons as assertions.
