---
title: "Reports"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
library(ksCompare)
```

A `ks_comparison` can be exported to two report formats. Both are
opt-in via `Suggests` packages so the core install stays small, and
neither requires Quarto, Pandoc, or any internet access.

## HTML

```{r}
cmp <- ks_compare(iris, iris, by = "Species")
ks_report_html(
  cmp,
  "report.html",
  title = "ADSL QC",
  subtitle = "Study A1234"
)
```

The HTML report uses `htmltools` for layout and `reactable` for
filterable, searchable tables. Layout includes:

- a sticky left-hand table of contents with section badges,
- a status pill (Clean / Differences detected) in the header,
- KPI cards summarising matched / unmatched rows and columns,
- Schema, Key/Row diff, Value diff, Patterns, Suggestions, and
  Manifest sections,
- a print-friendly stylesheet.

Pass `path = NULL` to get the assembled `htmltools::tagList()` for
embedding in another document. Pass `theme = "slate"` for a neutral
dark-header theme.

### Smart sampling

Large value-diff tables are capped at `max_rows` (default `2000`).
Above the cap, the report renders a *stratified sample* covering
each affected column and each distinct diff cause (`note`),
prioritising the largest numeric magnitudes. A notice indicates the
sample size and recommends `as_tibble(cmp)` for the full table
(`as_tibble()` is never truncated).

### Group by key

Pass `group_by_key = TRUE` to render the value-diff section as one
collapsed `<details>` block per key value, sorted by number of diffs
(most-affected first). Useful for clinical reviews where you want to
inspect "what changed for subject X" rather than scrolling a flat
table.

```{r}
ks_report_html(cmp, "report.html", group_by_key = TRUE, max_groups = 100)
```

When `dup_keys = "keep_all"` or `"all_pairs"` was used in
[ks_compare()], each key block also shows a *Pair* column so you can
tell duplicate-row pairings apart. If no `by =` was supplied (row-
position match), `group_by_key` is silently ignored.

## Excel

```{r}
ks_report_xlsx(cmp, "report.xlsx")
```

Sheets:

- `Summary` -- headline counts.
- `Schema` -- per-column metadata diff.
- `Suggestions` -- fuzzy column rename candidates (when
  `stringdist` is installed).
- `KeyDiff` -- matched / base-only / comp-only row counts.
- `Values` -- long cell-level diff table.
- `Patterns` -- detected recurring shapes.
- `OUT_BASE`, `OUT_COMP`, `OUT_DIF`, `OUT_NOEQUAL` -- PROC COMPARE-
  style wide outputs (see [as_outbase()] et al.).
- `Manifest` -- input hashes, package version, run timestamp.

Numeric `OUT_DIF` cells whose magnitude exceeds `threshold` (default
`0`) are highlighted via openxlsx2 conditional formatting. Pass
`highlight = FALSE` to suppress.
